Robinson et al. 2020 Supplemental Information: Comparing the Radiocarbon and Tree-ring Records for the Upland US Southwest

### 2020-03-11 analysis/paper/paper.Rmd
re-running everything

radiocarbon_spd <-
  all_data$Radiocarbon %>%
  split(., .$`Study Area`) %>%
  tibble::tibble(
    `Study Area` = names(.),
    Dates = .
  ) %>%
  dplyr::bind_rows(tibble::tibble(
    `Study Area` = "UUSW",
    Dates = list(all_data$Radiocarbon)
  )) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    Calibrated = list(Dates %$%
      rcarbon::calibrate(
        x = Date_BP,
        errors = SD,
        timeRange = timerange,
        calMatrix = TRUE,
        verbose = FALSE
      )),
    Bins = list(robinson2020::binPrep(
      sites = Dates$SiteID,
      ages = Dates$Date_BP,
      h = 100
    )),
    SPD = list(rcarbon::spd(
      x = Calibrated,
      bins = Bins,
      timeRange = timerange,
      spdnormalised = TRUE,
      verbose = FALSE
    ))
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(Density = SPD %>%
    purrr::map(function(x) {
      x$grid %>%
        tibble::as_tibble() %$%
        tibble::tibble(
          Type = "Radiocarbon Density",
          Date_BP = calBP,
          Density = PrDens
        )
    })) %T>%
  readr::write_rds(
    here::here("analysis/data/derived_data/radiocarbon_spd.rds"),
    compress = "gz"
  )
```

@@

dendro_spd <-
  all_data$`Tree-ring` %>%
  split(., .$`Study Area`) %>%
  tibble::tibble(
    `Study Area` = names(.),
    Dates = .
  ) %>%
  dplyr::bind_rows(tibble::tibble(
    `Study Area` = "UUSW",
    Dates = list(all_data$`Tree-ring`)
  )) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    Calibrated = list(Dates %$%
      calibrate_tree_ring_dates(
        x = Date_BP,
        timeRange = timerange
      )),
    Bins = list(robinson2020::binPrep(
      sites = Dates$SiteID,
      ages = Dates$Date_BP,
      h = 100
    )),
    SPD = list(rcarbon::spd(
      x = Calibrated,
      bins = Bins,
      timeRange = timerange,
      spdnormalised = TRUE,
      verbose = FALSE
    ))
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(Density = SPD %>%
    purrr::map(function(x) {
      x$grid %>%
        tibble::as_tibble() %$%
        tibble::tibble(
          Type = "Tree-ring Density",
          Date_BP = calBP,
          Density = PrDens
        )
    })) %T>%
  readr::write_rds(
    here::here("analysis/data/derived_data/dendro_spd.rds"),
    compress = "gz"
  )
```

@@

dplyr::left_join(all_density,
  vep_demography,
  by = c(
    "Study Area",
    "Year AD"
  )
) %>%
  dplyr::group_by(
    `Study Area`,
    Type
  ) %>%
  dplyr::mutate(`Scaled Density` = Density * (sum(Population, na.rm = T) / sum(Density * !is.na(Population), na.rm = TRUE))) %>%
  dplyr::group_by(
    `Study Area`,
    Type,
    Period
  ) %>%
  dplyr::group_by(
    `Study Area`,
    Type
  ) %>%
  dplyr::mutate(
    `Raw Density` = ifelse(`Study Area` == "UUSW", Density, `Scaled Density`),
    `Smoothed Density` = ifelse(Type == "Tree-ring Density", stats::filter(`Raw Density`, filter = dist), `Raw Density`),
    `Raw Density` = ifelse(`Study Area` == "UUSW", `Raw Density`, NA)
  ) %>%
  dplyr::select(
    `Study Area`,
    Type,
    `Year AD`,
    Population,
    `Raw Density`,
    `Smoothed Density`,
    # `Binned Density`
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    Type = factor(Type,
      levels = c(
        "Tree-ring Density",
        "Radiocarbon Density",
        "VEP Population"
      ),
      ordered = TRUE
    ),
    `Study Area` = factor(`Study Area`,
      levels = c(
        "UUSW",
        "CMV",
        "NRG"
      ),
      labels = c(
        "Upland US Southwest Density",
        "Central Mesa Verde Population",
        "Northern Rio Grande Population"
      ),
      ordered = TRUE
    )
  ) %>%
  {
    ggplot(
      .,
      aes(x = `Year AD`)
    ) +
      geom_line(aes(
        y = `Raw Density`,
        color = Type
      ),
      size = 0.5,
      alpha = 0.5
      ) +
      geom_line(aes(
        y = `Population`
      ),
      size = 1.25
      ) +
      geom_line(aes(
        y = `Smoothed Density`,
        color = Type
      ),
      size = 0.75
      ) +
      scale_color_manual(
        values = c(
          "Tree-ring Density" = "#d95f02",
          "Radiocarbon Density" = "#1b9e77",
          "VEP Population" = "black"
        ),
        limits = c(
          "Tree-ring Density",
          "Radiocarbon Density",
          "VEP Population"
        )
      ) +
      xlab("Year AD") +
      scale_x_continuous(
        limits = c(200, 1800),
        breaks = seq(200, 1800, 200),
        expand = expansion(0, 0)
      ) +
      theme_minimal(18) +
      theme(
        legend.title = element_blank(),
        legend.justification = c(0, 1),
        legend.position = "bottom",
        axis.title.x = element_text(size = 12),
        axis.text.x = element_text(size = 10),
        axis.title.y = element_blank(),
        strip.text = element_text(size = 12),
        # axis.text.y = element_blank(),
        # panel.grid.major.y = element_blank(),
        # panel.grid.minor.y = element_blank(),
        plot.margin = unit(c(0.25, 0.5, 0.25, 0.25), "in")
      ) +
      facet_wrap(
        # rows =
        vars(`Study Area`),
        ncol = 1,
        scales = "free_y"
      )
  } %>%
  egg::tag_facet(
    open = "",
    close = "",
    tag_pool = paste0(LETTERS, ". ", c(
      "Upland US Southwest Density",
      "Central Mesa Verde Population",
      "Northern Rio Grande Population"
    )),
    hjust = 0,
    vjust = 1.5,
    size = 5
  )
```

>>>>

radiocarbon_spd <-
  all_data$Radiocarbon %>%
  split(., .$`Study Area`) %>%
  tibble::tibble(
    `Study Area` = names(.),
    Dates = .
  ) %>%
  dplyr::bind_rows(tibble::tibble(
    `Study Area` = "UUSW",
    Dates = list(all_data$Radiocarbon)
  )) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    Calibrated = list(Dates %$%
      rcarbon::calibrate(
        x = Date_BP,
        errors = SD,
        timeRange = timerange,
        calMatrix = TRUE,
        verbose = FALSE
      )),
    Bins = list(robinson2020::binPrep(
      sites = Dates$SiteID,
      ages = Dates$Date_BP,
      h = 100,
      method = "single"
    )),
    SPD = list(rcarbon::spd(
      x = Calibrated,
      bins = Bins,
      timeRange = timerange,
      spdnormalised = TRUE,
      verbose = FALSE
    ))
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(Density = SPD %>%
    purrr::map(function(x) {
      x$grid %>%
        tibble::as_tibble() %$%
        tibble::tibble(
          Type = "Radiocarbon Density",
          Date_BP = calBP,
          Density = PrDens
        )
    })) %T>%
  readr::write_rds(
    here::here("analysis/data/derived_data/radiocarbon_spd.rds"),
    compress = "gz"
  )
```

@@

dendro_spd <-
  all_data$`Tree-ring` %>%
  split(., .$`Study Area`) %>%
  tibble::tibble(
    `Study Area` = names(.),
    Dates = .
  ) %>%
  dplyr::bind_rows(tibble::tibble(
    `Study Area` = "UUSW",
    Dates = list(all_data$`Tree-ring`)
  )) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    Calibrated = list(Dates %$%
      calibrate_tree_ring_dates(
        x = Date_BP,
        timeRange = timerange
      )),
    Bins = list(robinson2020::binPrep(
      sites = Dates$SiteID,
      ages = Dates$Date_BP,
      h = 100,
      method = "single"
    )),
    SPD = list(rcarbon::spd(
      x = Calibrated,
      bins = Bins,
      timeRange = timerange,
      spdnormalised = TRUE,
      verbose = FALSE
    ))
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(Density = SPD %>%
    purrr::map(function(x) {
      x$grid %>%
        tibble::as_tibble() %$%
        tibble::tibble(
          Type = "Tree-ring Density",
          Date_BP = calBP,
          Density = PrDens
        )
    })) %T>%
  readr::write_rds(
    here::here("analysis/data/derived_data/dendro_spd.rds"),
    compress = "gz"
  )
```

@@

dplyr::left_join(all_density,
  vep_demography,
  by = c(
    "Study Area",
    "Year AD"
  )
) %>%
  dplyr::group_by(
    `Study Area`,
    Type
  ) %>%
  dplyr::mutate(`Scaled Density` = Density * (sum(Population, na.rm = T) / sum(Density * !is.na(Population), na.rm = TRUE))) %>%
  dplyr::group_by(
    `Study Area`,
    Type,
    Period
  ) %>%
  dplyr::group_by(
    `Study Area`,
    Type
  ) %>%
  dplyr::mutate(
    `Raw Density` = ifelse(`Study Area` == "UUSW", Density, `Scaled Density`),
    `Smoothed Density` = ifelse(Type == "Tree-ring Density", stats::filter(`Raw Density`, filter = dist), `Raw Density`),
    `Raw Density` = ifelse(`Study Area` == "UUSW", `Raw Density`, NA)
  ) %>%
  dplyr::select(
    `Study Area`,
    Type,
    `Year AD`,
    Population,
    `Raw Density`,
    `Smoothed Density`,
    # `Binned Density`
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    Type = factor(Type,
      levels = c(
        "Tree-ring Density",
        "Radiocarbon Density",
        "VEP Population"
      ),
      ordered = TRUE
    ),
    `Study Area` = factor(`Study Area`,
      levels = c(
        "UUSW",
        "CMV",
        "NRG"
      ),
      labels = c(
        "Upland US Southwest Density",
        "Central Mesa Verde Population",
        "Northern Rio Grande Population"
      ),
      ordered = TRUE
    )
  ) %>%
  {
    ggplot(
      .,
      aes(x = `Year AD`)
    ) +
      geom_line(aes(
        y = `Raw Density`,
        color = Type
      ),
      size = 0.5,
      alpha = 0.5
      ) +
      geom_line(aes(
        y = `Population`
      ),
      size = 1.25
      ) +
      geom_line(aes(
        y = `Smoothed Density`,
        color = Type
      ),
      size = 0.75
      ) +
      scale_color_manual(
        values = c(
          "Tree-ring Density" = "#d95f02",
          "Radiocarbon Density" = "#1b9e77",
          "VEP Population" = "black"
        ),
        limits = c(
          "Tree-ring Density",
          "Radiocarbon Density",
          "VEP Population"
        )
      ) +
      xlab("Year AD") +
      scale_x_continuous(
        limits = c(200, 1800),
        breaks = seq(200, 1800, 200),
        expand = expansion(0, 0)
      ) +
      theme_minimal(18) +
      theme(
        legend.title = element_blank(),
        legend.justification = c(0, 1),
        legend.position = "bottom",
        legend.text = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        axis.text.x = element_text(size = 10),
        axis.title.y = element_blank(),
        axis.text.y = element_text(size = 10),
        # panel.grid.major.y = element_blank(),
        # panel.grid.minor.y = element_blank(),
        plot.margin = unit(c(0.25, 0.5, 0.25, 0.25), "in")
      ) +
      facet_wrap(
        # rows =
        vars(`Study Area`),
        ncol = 1,
        scales = "free_y"
      )
  } %>%
  egg::tag_facet(
    open = "",
    close = "",
    tag_pool = paste0(LETTERS, ". ", c(
      "Upland US Southwest Density",
      "Central Mesa Verde Population",
      "Northern Rio Grande Population"
    )),
    hjust = 0,
    vjust = 1.5,
    size = 4
  )
```
